---
title: "Lab 5: Logistic regression"
author: "Makesh Srinivasan"
date: "3/1/2022"
output: html_document
---

Registration number: 19BCE1717<br>
Faculty: Prof. Radha R<br>
Slot: L37 + L38<br>
Course code: CSE3505 <br>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Logistic regression

***

<strong id="part1">Logistic regression</strong>: Admissions datset<br>
<br>
Use the dataset <strong>admission.csv</strong> to train a Logistic regression model
<br>
<br>

##### Libraries:
```{r}
library(caret)
library(ggplot2)
```
<br>

##### Dataset
```{r}
rm(list=ls())
mydata <- read.csv('admission.csv',header=T)
```
<br>

##### Dataset exloration

Structure:
```{r}
str(mydata)
```

Head:
```{r}
head(mydata)
```

Summary:
```{r}
summary(mydata)
```

Histogram plot:
```{r}
hist(mydata$admit,col='blue')
```


```{r}
mydata$admit <- as.factor(mydata$admit)
summary(mydata)
```
<br>

##### Training 
Loading the library and splitting 80-20 Training-Testing
```{r}
train_ind=createDataPartition(mydata$admit,p=0.80,list=F)
train_ds=mydata[train_ind,]
test_ds=mydata[-train_ind,]
```

Training the model:
```{r}
my_model=glm(admit~.,data=train_ds,family = "binomial")
summary(my_model)
```
<br>

##### Predictions
Running predictions on test-set:
```{r}
pred_train=predict(my_model,train_ds,type="response")
pred_train
```
<br>

##### Performance analysis

<strong>Train set</strong><br>
Confusion matrix:
```{r}
pred_train=ifelse(pred_train>0.5,1,0)
conf <- table(Predicted=pred_train,Actual=train_ds$admit)
conf
```

Misclassification error:
```{r}
1-sum(diag(conf))/sum(conf)
```

<strong>Test set</strong><br>
Confusion matrix:
```{r}
pred_test <- predict(my_model, test_ds, type = 'response')
pred_test <- ifelse(pred_test>0.5, 1, 0)
tab2 <- table(Predicted = pred_test, Actual = test_ds$admit)
tab2
```

Misclassification error:
```{r}
1 - sum(diag(tab2))/sum(tab2)
```

***

