---
title: 'Lab 10: Random forest'
author: "Makesh Srinivasan"
date: "4/13/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Registration number: 19BCE1717<br>
Faculty: Prof. Radha R<br>
Slot: L37 + L38<br>
Course code: CSE3505 <br>

***


## Random forest

***

## Sections:
* <a href="#part1">Dataset 1 - Seeds</a>
* <a href="#part2">Dataset 2 - Crabs</a>

***

<center><h3 id="part1">Seeds dataset</h3><br></center>
<br>

<br>

##### Load necessary libraries:

```{r}
library('stats19')
library('dplyr')
library('randomForest')
```
<br>

##### Load the dataset:

```{r}
mydata = read.csv("Seeds.csv")
str(mydata)
```
<br>

##### Split 70:30 train-test:

```{r}
mydata$seedType = as.factor(mydata$seedType)
index=sample(2,nrow(mydata), replace=TRUE,prob=c(0.7,0.3))
training=mydata[index==1,]
testing=mydata[index==2,]
```
<br>

##### Train the random forest model:

```{r}
RFM <- randomForest(seedType ~ .,data=training, importance=T, proximity=T)
seed_pred = predict(RFM,testing)
testing$seed_pred = seed_pred
head(testing)
```

<br>

##### Confusion matrix:
```{r}
CFM = table(testing$seedType,testing$seed_pred)
CFM
```


<br>

##### Accuracy:
```{r}
acc = sum(diag(CFM))/sum(CFM)*100
acc
```

<br>
<strong>INFERENCE: </strong>We see that the accuracy is 100 which means all data points are accurately classified. This was also observed in the previous code chunk - confusion matrix. There were no misclassification of see types. <br><br>

Now let us see another dataset.

***

<center><h3 id="part2">Crabs dataset</h3><br></center>
<br>


##### Crabs dataset
```{r}
library(MASS)
crabs <- crabs
str(crabs)
```

<br>

##### Random forest for species on crabs:

```{r}
index=sample(2,nrow(crabs), replace=TRUE,prob=c(0.5,0.5))
training=crabs[index==1,]
testing=crabs[index==2,]

RFM <- randomForest(sp ~ .,data=training, importance=T, proximity=T)
Species_Pred = predict(RFM, testing)
testing$Species_Pred=Species_Pred
```

<br>

##### Performance measure:

```{r}
# confusion matrix
CFM = table(testing$sp,testing$Species_Pred)
CFM

# accuracy
acc = sum(diag(CFM))/sum(CFM)*100
acc
```

<br>
<strong>INFERENCE: </strong>We see that the accuracy is 89.19 which means most of the data points are accurately classified. This was also observed in the confusion matrix. Only 12 misclassifications occurred - 10 Blue were incorrectly predicted as Orange while 2 Orange were incorrectly predicted as Blue.

***


