---
title: "Lab 4: One-way ANOVA"
author: "Makesh Srinivasan"
date: "2/22/2022"
output: html_document
---

Registration number: 19BCE1717<br>
Faculty: Prof. Radha R<br>
Slot: L37 + L38<br>
Course code: CSE3505 <br>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## One-way ANOVA

***

## Sections:
* <a href="#part1"><strong>Part 1: </strong>Machines dataset</a>
* <a href="#part2"><strong>Part 2: </strong>Football dataset (lab exercise)</a>

***
<center><h3 id="part1">Part 1</h3></center>
<center><strong>Machines dataset</strong><br></center>
Use the dataset <strong>onewayanova_machine.csv</strong> to perform one-way ANOVA
<br>
<br>

##### Libraries:
```{r}
library(ggplot2)
```

##### Dataset
```{r}
combined_groups <- read.csv("onewayanova_machine.csv")
combined_groups
```
<br>
Summary statistics:
```{r}
summary(combined_groups)
```
<br>
Boxplot:
```{r}
boxplot(combined_groups,ylab="Diameter",xlab="Machines",col="orange")
```

Editing the dataframe to suitable structure:
```{r}
stacked_groups <- stack(combined_groups)
stacked_groups
```
<br>
Rename the attributes:
```{r}
names(stacked_groups) <- c("Diameter","Machines")
stacked_groups
```
<br>

##### Performing ANOVA and HSD test

Performing one way anova:
```{r}
anova_results <- aov(Diameter~Machines,data=stacked_groups)
summary(anova_results)
```
<br>
Performing Tuckey HSD test on the ANOVA results:
```{r}
tk <- TukeyHSD(anova_results)
tk
```
<br>

##### Visualise

Visualisation of the means of the Machines:
```{r}
plot(tk,col="blue")
```

***
<center><h3 id="part2">Part 2</h3></center>
<center><strong>Football dataset</strong></center>
<br>
Use the dataset <strong>One way Anova_Football.csv</strong> to perform one-way ANOVA
<br>
<br>

##### Dataset
```{r}
combined_groups <- read.csv("One way Anova_Football.csv")
combined_groups
```

<br>
Summary statistics:
```{r}
summary(combined_groups)
```
<br>
Boxplot:
```{r}
boxplot(combined_groups,ylab="Weight",xlab="Teams/clubs",col="#984447")
```
<br>
Editing the dataframe to suitable structure:
```{r}
stacked_groups <- stack(combined_groups)
stacked_groups
```
<br>
Rename the attributes:
```{r}
names(stacked_groups) <- c("Weight","Clubs")
stacked_groups
```

<br>

##### Performing ANOVA and HSD test

Performing one way anova:
```{r}
anova_results <- aov(Weight~Clubs,data=stacked_groups)
summary(anova_results)
```
<br>
Performing Tuckey HSD test on the ANOVA results:
```{r}
tk <- TukeyHSD(anova_results)
tk
```
<br>

##### Visualise

Visualisation of the means of the Weights for clubs:
```{r}
plot(tk,col="blue")
```

***
