---
title: "Lab 1: Linear regression"
author: "Makesh Srinivasan"
date: "1/11/2022"
output: html_document
---

Registration number: 19BCE1717<br>
Faculty: Prof. Radha R<br>
Slot: L37 + L38<br>
Course code: CSE3505 <br>

***

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Linear regression

From the dataset given in class, generate a linear model and explore the same. Exercise question: Predict for when x = 1200 using the linear model

##### 1) Create the dataset
Declare and initialise the columns of the dataframe:
```{r}
x <- c(200, 500, 700, 1000)
y <- c(0.25, 0.5, 0.8, 1.5)
```

Form the dataframe:
```{r}
data <- data.frame(x,y)
# Uncomment the following line to view the data. While knitting, the View() function does not compile. Thus, I have used head() 
# View(data)
head(data) # Display the dataset
```

##### 2) Explore the data

Structure of the data:
```{r}
str(data)
```

Attrbutes in the dataset:
```{r}
names(data)
```

Visualise the data
```{r}
plot(x,y)
```

##### 3) Linear regression model

Create a linear model with hypothesis y = w0 + w1x:
```{r}
model_1 <- lm(y~x)
model_1
```

Summary of the model:
```{r}
summary(model_1)
```

Construct linear line on the data points:
```{r}
plot(x,y)
abline(model_1, col="blue")
```

##### 4) Testing the regressed model:
<b>Approach 1:</b>
```{r}
test_x1 <- data.frame(x=6:8)
R <- predict(model_1, test_x1)
R #predicted values
```

Summary of test data:
```{r}
summary(R)
```

<b>Approach 2:</b>
```{r}
test_x2 <- data.frame(x=c(6,7,8))
R1 <- predict(model_1, test_x2)
R1
```

Summary of test data:
```{r}
summary(R1)
```

***
##### 5) Exercise question
1) Find y if x = 1200
```{r}
x <- data.frame(x = 1200)
R2 <- predict(model_1, x)
R2
```

***
