---
title: 'Lab 9: Hierachical clustering'
author: "Makesh Srinivasan"
date: "4/5/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Registration number: 19BCE1717<br>
Faculty: Prof. Radha R<br>
Slot: L37 + L38<br>
Course code: CSE3505 <br>

***


## Hierachical clustering
##### Complete, average and single linkage

***

## Sections:
* <a href="#part1">Class task - USArrests</a>
* <a href="#part2">Seeds dataset</a>

***

<center><h3 id="part1">Part 1</h3><br>
Class task - USArrests</center>
<br>

<br>

##### Load necessary libraries
```{r}
rm(list=ls())
#install.packages("factoextra")
#install.packages("cluster")
library(factoextra)
library(cluster)
```
<br>

##### Load and preparing the data

```{r}
df <- USArrests  
df <- na.omit(df)
head(df)
df <- scale(df) # scaling to normalize the values >>>> x-xbar/SD
head(df)
```

<br>

##### Hierachical clustering "Complete linkage"

Show the cut at clusters = 4
```{r}
dist_mat <- dist(df, method = 'euclidean')
hclust_complete <- hclust(dist_mat, method = 'complete') # function call for Hier. Clust.
plot(hclust_complete, cex = 0.5, hang = -1)
# hang = A negative value will cause the labels to hang down from 0.
# cex= font size
cut<- cutree(hclust_complete, k = 4)
abline(h = 4, col = 'red')
rect.hclust(hclust_complete, k = 4, border = 2:5)
```

<br>

##### Hierachical clustering "Average linkage"

Show the cut at clusters = 4
```{r}
hclust_average <- hclust(dist_mat, method = 'average')
plot(hclust_average, cex = 0.5, hang = -1)
cut<- cutree(hclust_average, k = 4)
abline(h = 2.5, col = 'red')
rect.hclust(hclust_average, k = 4, border = 2:5)
```
<br>

##### Hierachical clustering "Single linkage"

Show the cut at clusters = 4
```{r}
hclust_simple <- hclust(dist_mat, method = 'single')
plot(hclust_simple, cex = 0.5, hang = -1)
cut<- cutree(hclust_simple, k = 4)
abline(h = 1.3, col = 'red')
rect.hclust(hclust_simple, k = 4, border = 2:5)
```
<br>
NOTE: The line cuts the dendrogram at four clusters but it is not very clear as the clustering is too dense
<br>

***

<center><h3 id="part2">Part 2</h3><br>
Seeds dataset</center>
<br>

```{r}
df <- read.csv("seeds_K Means.csv")
df<-na.omit(df)
df<-scale(df) # Normalise the dataset
head(df)
```


<br>

##### Hierachical clustering "Complete linkage"

Show the cut at clusters = 3
```{r}
dist_mat <- dist(df, method = 'euclidean')
hclust_complete <- hclust(dist_mat, method = 'complete') # function call for Hier. Clust.
plot(hclust_complete, cex = 0.2, hang = -1)
# hang = A negative value will cause the labels to hang down from 0.
# cex= font size
cut<- cutree(hclust_complete, k = 3)
abline(h = 6.2, col = 'red')
rect.hclust(hclust_complete, k = 3, border = 2:5)
```


<br>

##### Hierachical clustering "Average linkage"

Show the cut at clusters = 3
```{r}
hclust_average <- hclust(dist_mat, method = 'average')
plot(hclust_average, cex = 0.2, hang = -1)
cut<- cutree(hclust_average, k = 3)
abline(h = 4, col = 'red')
rect.hclust(hclust_average, k = 3, border = 2:5)
```
<br>

##### Hierachical clustering "Single linkage"

Show the cut at clusters = 3
```{r}
hclust_simple <- hclust(dist_mat, method = 'single')
plot(hclust_simple, cex = 0.2, hang = -1)
cut<- cutree(hclust_simple, k = 3)
abline(h = 1.7, col = 'red')
rect.hclust(hclust_simple, k = 3, border = 2:5)
```
<br>
NOTE: The line overlaps with the thre box borders and I did not move it further up because then it will cut the dendrogram at k = 2
