---
title: "Lab 3: Time series forecasting - 2 (Non stationary)"
author: "Makesh Srinivasan"
date: "1/25/2022"
output: html_document
---

Registration number: 19BCE1717<br>
Faculty: Prof. Radha R<br>
Slot: L37 + L38<br>
Course code: CSE3505 <br>

***

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Time series forecasting 2 - Non stationary
<strong> Dataset: </strong>soya_oil data set for TS forecast.csv
<br>
<br>


##### Import necessary libraries
```{r}
rm(list=ls())
library('forecast')
library('tseries')
```
<br>

##### Dataset
```{r}
data=read.csv("soya_oil data set for TS forecast.csv")
class(data) # dataframe
```

Convert the data to timeseries:
```{r}
data=ts(data$Dollar,start=1,end=128,frequency=4)
class(data) # time series
# View(data)
```
<br>

##### Visualisations

Visualise the timeseries data:
```{r}
plot(data,xlab='time',ylab='Soya Oil sales')
```

Visualizing diff(data):
```{r}
dd <- diff(data)
plot(dd,ylab='Differenced Soya Oil Sales')
```

Visualizing log(data):
```{r}
ld <- log10(data)
plot(ld,ylab='Differenced Log (Soya Oil Sales)')
```

Visualizing diff(log(data)):
```{r}
ld <- log10(data)
dld <- diff(ld)
plot(dld,ylab='Differenced Log (Soya Oil Sales)')
```
<br>


##### Testing and plotting
```{r}
adf.test(data)
```

ACF and PACF plots + fitting ARIMA model
```{r}
par(mfrow=c(1,2))
acf(ts(data),main="ACF Soya Oil Sales")
pacf(ts(data),main="PACF Soya Oil Sales")
```

```{r}
ARIMAfit=auto.arima(data,ic="aic",trace=TRUE)
```

```{r}
summary(ARIMAfit)
```

Forecasting 3 time periods with the fitted model
```{r}
fc=forecast(ARIMAfit,level=c(95),h=3*12)
plot(fc)
```

```{r}
summary(fc)
```

```{r}
attributes(fc)
```

***
