---
title: "Lab 2: Time series forecasting 1"
author: "Makesh Srinivasan"
date: "1/25/2022"
output: html_document
---

Registration number: 19BCE1717<br>
Faculty: Prof. Radha R<br>
Slot: L37 + L38<br>
Course code: CSE3505 <br>

***

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Time series forecasting 1
<strong> Dataset: </strong>Tractor-sales.csv
<br>
<br>


##### Import necessary libraries
```{r}
rm(list=ls())
library('forecast')
library('tseries')
```
<br>

##### Dataset
```{r}
data <- read.csv("Tractor-sales.csv")
class(data) # this is now a dataframe
```

Convert the data to timeseries:
```{r}
data=ts(data$Number.of.Tractor.Sold,start=c(2003,1),frequency=12)
class(data) # This is now a time series dataset
```
<br>

##### Visualisations

Visualise the timeseries data:
```{r}
plot(data,xlab='Years',ylab='Tractor sales')
```

Visualizing diff(data):
```{r}
plot(diff(data),ylab='Differenced Tractor Sales')
```

Visualizing log(data):
```{r}
ld <- log10(data)
plot(ld,ylab='Log (Tractor sales)')
```

Visualizing diff(log(data)):
```{r}
dld <- diff(ld)
plot(dld,ylab='Differenced Log (Tractor Sales)')
```
<br>

##### Testing and plotting
```{r}
adf.test(dld)
```

ACF and PACF plots + fitting ARIMA model
```{r}
par(mfrow=c(1,2))
acf(ts(dld),main="ACF Tractor sales")
pacf(ts(dld),main="PACF Tractor sales")
```

```{r}
ARIMAfit=auto.arima(dld,ic="aic",trace=TRUE)
```

```{r}
summary(ARIMAfit)
```

Forecasting 3 time periods with the fitted model
```{r}
fc=forecast(ARIMAfit,level=c(95),h=3*12)
plot(fc)
```

```{r}
summary(fc)
```

***
