---
title: "Lab 6: K-Means clustering"
author: "Makesh Srinivasan"
date: "3/8/2022"
output: html_document
---

Registration number: 19BCE1717<br>
Faculty: Prof. Radha R<br>
Slot: L37 + L38<br>
Course code: CSE3505 <br>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## K-Means clustering

***

## Sections:
* <a href="#part1">USArrests</a>
* <a href="#part2">Seeds K Means</a>

***
<center><h3 id="part1">Part 1: USArrests</h3></center>
<center><strong>K-Means clustering</strong>: USArrests</center><br>
<br>
Use the built-in dataset <strong>USArrests</strong> to perform clustering using K-Means
<br>
<br>

##### Libraries:
```{r}
library(factoextra)
library(cluster)
```
<br>

##### Dataset
```{r}
df <- USArrests
df<-na.omit(df)
df<-scale(df) # Normalise the dataset
head(df)
```

<br>

##### Optimal number of clusters

Plot number of clusters vs total within sum of squares
```{r}
fviz_nbclust(df,kmeans,method="wss")
```
<br>We notice that after scaling, we get a bend at k=4 so that is the optimal number of clusters

<br>

##### Clustering 

Method: K-Means <br>
k = 4 <br>
```{r}
set.seed(1) #to reproduce same results always
km <- kmeans(df, centers = 4, nstart = 25)
km
```

Plot the results of the clustering:
```{r}
fviz_cluster(km, data=df)
```

<br>

Add the clusters assigned to the dataset:
```{r}
final_data<-cbind(USArrests,cluster=km$cluster)
```

<br>

##### Clustered dataset (final)

```{r}
head(final_data)
```

***
<center><h3 id="part2">Part 2: Seeds</h3></center>
<center><strong>K-Means clustering</strong>: seeds_K Means.csv </center><br>
<br>
Use the dataset <strong>seeds_K Means.csv</strong> to perform clustering using K-Means
<br>
<br>

##### Dataset
```{r}
df <- read.csv("seeds_K Means.csv")
df<-na.omit(df)
df<-scale(df) # Normalise the dataset
head(df)
```

<br>

##### Optimal number of clusters

Plot number of clusters vs total within sum of squares
```{r}
fviz_nbclust(df,kmeans,method="wss")
```
<br>Same as the above scenario, we notice that after scaling, we get a bend at k=4 so that is the optimal number of clusters

<br>

##### Clustering 

Method: K-Means <br>
k = 4 <br>
```{r}
set.seed(1) #to reproduce same results always
km <- kmeans(df, centers = 4, nstart = 25)
km
```

Plot the results of the clustering:
```{r}
fviz_cluster(km, data=df)
```

<br>

Add the clusters assigned to the dataset:
```{r}
final_data<-cbind(df,cluster=km$cluster)
```

<br>

##### Clustered dataset (final)

```{r}
head(final_data)
```

***


